%Bibliographic references
\begin{thebibliography}{9}


\bibitem{FHEW} \label{fhew1}
Léo Ducas and Daniele Micciancio,"FHEW: Bootstrapping Homomorphic Encryption in less than a second"
\textit{Cryptology ePrint Archive 2014/816,} 2014.

\bibitem{easyfhe} \label{easyfhe}
Craig Gentry,"Computing Arbitrary Functions of Encrypted Data"
\textit{https://crypto.stanford.edu/craig/easy-fhe.pdf} 

\bibitem{FHEW Library} \label{fhewlib}
L. Ducas and D. Micciancio. Implementation of FHEW.
\textit{https://github.com/lducas/FHEW} June 2014.

\bibitem{FFTW} \label{fftw}
Matteo Frigo and Steven G. Johnson,"Fastest Fourier Transform in the West"
\textit{http://www.fftw.org}.

\bibitem{FFTW_paper} \label{fftw_paper}
Matteo Frigo,"A Fast Fourier Transform Compiler"
\textit{MIT Laboratory for Computer Science}.February 1999.

\bibitem{Libfi} \label{lib}
Gabi Sarkis. 
\textit{https://github.com/gsarkis/libfi}. 

\bibitem{FHEW ideal Lattices} \label{lattices}
Craig Gentry,"Fully Homomorphic Encryption Using Ideal Lattices".
\textit{ In M. Mitzenmacher, editor,41st ACM STOC,pages 169–178. ACM Press} May / June 2009.

\bibitem{FFTW} \label{homo}
Brian Hayes,"Alice and Bob in Cipherspace"
\textit{http://www.americanscientist.org/libraries/documents/201286159329266-2012-09CompSciHayes.pdf}.

\bibitem{FFTW3} \label{FFTW3git}
"FFTW/fftw3"
\textit{https://github.com/FFTW/fftw3.git}.


\bibitem{FFT accuracy}  \label{FFT accuracy}
James C. Schatzman. Accuracy of the Discrete Fourier Transform and the Fast Fourier Transform.
\textit{https://pdfs.semanticscholar.org/9939/893e66a57e89271c5bcd2d64b2a961cf4425.pdf}

\bibitem{Papaa} \label{papaa}
Gopalakrishna Hegde, Nachiappan Ramasamy, Nachiket Kapre.OpenCL Labs for PAPAA Summer School 2016 Edition.
\textit{https://github.com/gplhegde/papaa-opencl}. 

\bibitem{Eli Hughes} 
Eli Hughes. (2014, April 28). Introduction to Fixed Point Math for Embedded Systems [Video File].Retrieved from
\textit{https://www.youtube.com/user/emh203/videos}

\bibitem{Double Precision} \label{dp}
Double-precision floating-point format.
\textit{https://en.wikipedia.org/wiki/Double-precision\_floating-point\_format}

\bibitem{Vivado Design Suite} \label{Vivado}
\textit{Vivado Design Suite User Guide, High-Level Synthesis}

\bibitem{accuracy fftw3} \label{accuracy fftw3}
benchFFT,\textit{http://www.fftw.org/accuracy/}

\bibitem{oppenheim} \label{accuracy fftw3}
Discrete-Time Signal Processing, Alan V. Oppenheim, Ronald W. Schafer,John R. Busk.
\textit{Prentice Hall,New Jeresy} 1998.

\bibitem{architechture} \label{architechture}
Architechture Exploration of FPGA based Accelerators for Bioinformatics Applications, B. Sharat Chandra Varma,Kolin Paul, M. Balakrishnan.
\textit{Springer Series in Advanced Microelectronics,Singapore} 2016.
\end{thebibliography}



%\end{document}


